---
title: "HR_analytics"
author: "Jiao Lai"
date: "4/4/2021"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Steps to analyze HR data:
1. Identify groups to compare.
2. Calculate summary statistics for each group.
3. Compare the differences statistically or visually.

Case overview:
1. Identifying the best recruiting source.
Quality of hire:
  retention, or how long the employee stays
  manager's satisfaction with the hire
  job performance
  amount of time it takes to become fully productive
2. What is driving low employee engagement?
3. Are new hires getting paid too much?
4. Are performance ratings being given consistently?
5. Improving employee safety with data.

Load data from website.
```{r}
library(readr)
library(broom)
survey <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/survey_data.csv"))
recruitment <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/recruitment_data.csv"))
pay <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/fair_pay_data.csv"))
performance <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/performance_data.csv"))
hr_1 <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/hr_data.csv"))
accident <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/accident_data.csv"))
hr_2 <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/hr_data_2.csv"))
survey_2 <- read_csv(url("https://assets.datacamp.com/production/course_5977/datasets/survey_data_2.csv"))
```
The dataset 'recruitment' contains sources of recruiting, and three measurements of quality of hires.
For this dataset, we are interested in whether quality of hires are different in terms of recruiting source.

```{r}
head(recruitment)
names(recruitment)
summary(recruitment)
colSums(is.na(recruitment))
levels(recruitment$recruiting_source)
library(dplyr)
library(tidyr)
recruitment_summary <- recruitment %>%
  group_by(recruiting_source) %>%
  summarize(
    count = n(),
    avg_attrition = mean(attrition),
    avg_rating = mean(performance_rating),
    avg_sales = mean(sales_quota_pct),
  ) %>%
  gather( "measurement", 'value', -recruiting_source)
library(ggplot2)
recruitment_summary %>%
  ggplot(aes(x = recruiting_source, y = value, fill = measurement))+
  geom_col(position = 'dodge') +
  facet_wrap(~measurement, scales = 'free')
```
The dataset 'survey' contains info for employees: salary, department, engagement, vacation days taken.
```{r}
head(survey)
summary(survey)
unique(survey$department)
survey_summary <- survey %>%
  mutate(disengaged = ifelse(engagement %in% c(1, 2), 1, 0)) %>%
  group_by(department) %>%
  summarize(
    count = n(),
    avg_engagement = mean(engagement),
    avg_salary = mean(salary),
    avg_vacation_days_taken = mean(vacation_days_taken),
    avg_disengaged = mean(disengaged)
  )
survey_summary %>%
  gather("measurement", "value", -department) %>%
  ggplot(aes(x = measurement, y = value, fill = department)) +
  geom_col(position = 'dodge') +
  facet_wrap(~measurement, scales = 'free')

```
To test if two groups are statistically significant different from each other, we can use t-test if the variable we are comparing is continuous, and chi-square test if the variable we are comparing is categorical.
```{r}
survey <- survey %>%
  mutate(in_sales = ifelse(department == 'Sales', "Sales", "Other"),
         disengaged = ifelse(engagement %in% c(1,2), 1, 0)) 
## check if Sales and other department have different 'disengaged' and 'vacation_days_taken'
chisq.test(survey$in_sales, survey$disengaged)
## check if the two groups have equal variance 
var.test(vacation_days_taken ~ in_sales, survey)
## check if the two groups are normal
with(survey, shapiro.test(vacation_days_taken[in_sales == 'Sales']))
with(survey, shapiro.test(vacation_days_taken[in_sales == 'Other']))
t.test(vacation_days_taken ~ in_sales, survey, var.equal = FALSE)
 
```

The table 'pay' contains employee_id, department, salary, whether they are new_hire, and their job_level.
```{r}
names(pay)
head(pay)
summary(pay)
pay %>%
  group_by(new_hire) %>%
  summarize(
    count = n(),
    avg_salary = mean(salary))
var.test(salary ~ new_hire, pay)
with(pay, shapiro.test(salary[new_hire == 'Yes']))
with(pay, shapiro.test(salary[new_hire == 'No']))
t.test(salary ~ new_hire, pay, var.equal = TRUE) 
```
Check for omitted variables.
```{r}
pay %>%
  ggplot(aes(x = new_hire, fill = job_level)) +
  geom_bar(position = 'fill')
pay %>%
  group_by(new_hire, job_level) %>%
  summarize(avg_salary = mean(salary)) %>%
  ggplot(aes(x = new_hire, y = avg_salary, fill = job_level)) +
  geom_col(position = 'dodge') +
  facet_wrap(~job_level, scales = 'free')

```
Look at 'hourly' job_level only.
```{r}
hourly <- pay %>% filter(job_level == 'Hourly') 
var.test(salary ~ new_hire, hourly)
tidy(t.test(salary ~ new_hire, hourly, var.equal = TRUE))

```
Use linear regression to control confounding variables.
```{r}
## Simple linear regression.
## Simple linear regression gives the same result with t test of equal variances.
lm.simple <- lm(salary ~ new_hire, pay) %>% tidy()
## Add job_level.
lm.mul <- lm(salary ~ new_hire + job_level, pay) %>% tidy()

```
Analyze HR data from different resources.
```{r}
summary(hr_1)
summary(performance)
joined <- hr_1 %>%
  left_join(performance, by = 'employee_id')
summary(joined)
joined %>%
  group_by(gender) %>%
  summarize(avg_rating = mean(rating, na.rm = TRUE))
```
Compare performance by gender.
```{r}
joined <- joined %>%
  mutate(high_performer = ifelse(rating >= 4, 1, 0))
## Compare the difference in performance by gender
chisq.test(joined$high_performer, joined$gender)
chisq.test(joined$rating, joined$gender)
## visualize the distribution of performance by gender
joined %>%
  ggplot(aes(x = gender, fill = factor(high_performer)))+
  geom_bar(position = 'fill')
joined %>%
  ggplot(aes(x = gender, fill = factor(rating)))+
  geom_bar(position = 'fill')
joined %>%
  ggplot(aes(x = gender, fill = factor(job_level)))+
  geom_bar(position = 'fill')
## check if job_level distribution is different by gender
chisq.test(joined$job_level, joined$gender)
## visualize the difference in performance by gender and job_level
joined %>%
  ggplot(aes(x = gender, fill = factor(high_performer)))+
  geom_bar(position = 'fill') +
  facet_wrap(~job_level, scales = 'free')
```
Use logistic regression to predict binary variable.
Find variables affecting an employee's chance to be high_performer.

```{r}
glm.simple <- glm(high_performer ~ gender, joined, family = 'binomial') %>% tidy()
glm.mul <- glm(high_performer ~ gender + job_level, joined, family = 'binomial')  %>% tidy()
```

Analyze workforce safety.
```{r}
head(hr_2)
head(accident)
acc_joined <- hr_2 %>%
  left_join(accident, by = c('employee_id', 'year')) %>%
  mutate(had_accident = ifelse(is.na(accident_type), 0, 1))
## Accident rate by year.
acc_joined %>%
  group_by(year) %>%
  summarize(avg_accident = mean(had_accident))
chisq.test(acc_joined$had_accident, acc_joined$year)
acc_joined %>%
  group_by(year, location) %>%
  summarize(avg_accident = mean(had_accident)) %>%
  ggplot(aes(x = factor(year), y = avg_accident, fill = location)) +
  geom_col(position = 'dodge') +
  facet_wrap(~location, scales = 'free')
```
Looked at subset of data with interest: Southfield.

```{r}
southfield <- acc_joined %>%
  filter(location == 'Southfield') 
southfield %>%
  group_by(year) %>%
  summarize(avg_overtime_hours_worked = mean(overtime_hours)) %>%
  ggplot(aes(x = year, y = avg_overtime_hours_worked)) +
  geom_col()
t.test(overtime_hours ~ year, southfield)
```
Use more data to check for sources of variation.
```{r}
acc_survey <- southfield %>% left_join(survey_2, by = c('employee_id', 'year')) %>%
  mutate(disengaged = ifelse(engagement <= 2, 1, 0),
         year = as.factor(year))
acc_survey %>%
  ggplot(aes(x = year, fill = factor(disengaged))) +
  geom_bar(position = 'fill')
chisq.test(acc_survey$disengaged, acc_survey$year)
## check accident trend in other fields.
other <- acc_joined %>% filter(location != 'Southfield') 
other %>%
  group_by(year) %>%
  summarize(avg_accident = mean(had_accident))
chisq.test(other$had_accident, other$year)
```
Use regression to control other variables.
```{r}
glm(had_accident ~ year + disengaged, data = acc_survey, family = 'binomial') %>% tidy()

```
